Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit

### 2023-06-16 Manuscript/Manuscript.Rmd
edit MS

To identify NOS, we obtained a "seed" database of oxygenase domain in Pfam database, PF02898.
From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
HMM models were run in HMMR3 with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.

@@

To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211.
From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
HMM models were run in HMMR3 with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster.
To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.

@@

We used Achim et al. for the single-cell data [@Achim2015].
In Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
Since the raw data were read count data, we normalized them to TPM using Python.
After that we converted them to log10.
From the sum of the expression levels in 107 cells for each gene, We calculated the percentage in each cell.
For each cell, we identified them with marker genes.
After created the data in Python, plotted it using R dot plots.
RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
In this case we wanted to compare between samples, so we used TPM.
The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.

>>>>

To identify NOS, we obtained a "seed" database of oxygenase domain in Pfam database, PF02898.
From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
HMM models were run in HMMR3 with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
The sequences used for the phylogenetic analysis are available in **Supplementary file 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.

@@

To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211.
From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
HMM models were run in HMMR3 with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster.
To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
The sequences used for the phylogenetic analysis are available in **Supplementary file 4**, the trimmed alignment is available in **Supplementary file 5** and the pre-trimmed data in **Supplementary file 6**.

@@

We used Achim et al. for the single-cell data [@Achim2015].
In Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
Since the raw data were read count data, we normalized them to TPM using Python.
After that we converted them to log10.
From the sum of the expression levels in 107 cells for each gene, We calculated the percentage in each cell.
For each cell, we identified them with marker genes.
After created the data in Python, plotted it using R dot plots.
RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
In this case we wanted to compare between samples, so we used TPM.
The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.
